---
title: "hw01-daniel-john"
author: "Daniel John"
date: "February 20, 2018"
output: github_document
---

## 2) Data Import
```{r}
column_names <- c('symboling', 
                  'normailzed_losses', 
                  'make', 
                  'fuel_type', 
                  'aspiration', 
                  'num_of_doors', 
                  'body_style',
                  'drive_wheels',
                  'engine_location',
                  'wheel_base',
                  'length',
                  'width',
                  'height',
                  'curb_weight',
                  'engine_type',
                  'num_of_cylinders',
                  'engine_size',
                  'fuel_system',
                  'bore',
                  'stroke',
                  'compression_ratio',
                  'horsepower',
                  'peak_rpm',
                  'city_mpg',
                  'highway_mpg',
                  'price')
```

```{r}

library(readr)

column_classes <- cols(
  'symboling' = col_double(),
  'normalized_losses' = col_double(),
  'make' = col_character(),
  'fuel_type' = col_character(),
  'aspiration' = col_character(),
  'num_of_doors' = col_character(),
  'body_style' = col_character(),
  'drive_wheels' = col_character(),
  'engine_location' = col_character(),
  'wheel_base' = col_double(),
  'length' = col_double(),
  'width' = col_double(),
  'height' = col_double(),
  'curb_weight' = col_integer(),
  'engine_type' = col_character(),
  'num_of_cylinders' = col_character(),
  'engine_size' = col_integer(),
  'fuel_system' = col_character(),
  'bore' = col_double(),
  'stroke' = col_double(),
  'compression_ratio' = col_double(),
  'horsepower' = col_integer(),
  'peak_rpm' = col_integer(),
  'city_mpg' = col_integer(),
  'highway_mpg' = col_integer(),
  'price' = col_integer()
)
```

```{r}
imports_85_table <- read.csv(file = 'imports_85.data', 
         header = FALSE, 
         col.names = column_names,
         na.strings = '?',
         colClasses = column_classes
         )
imports_85_table

str(imports_85_table)

object.size(imports_85_table)
```

```{r}
library(readr)

imports_85_table_readr <- read_csv('imports_85.data',
         col_names = column_names,
         col_types = column_classes,
         na = c("?","NA")
         )
imports_85_table_readr

object.size(imports_85_table)

str(imports_85_table_readr)
```

## 3) Technical Questions about importing data
# a. Simply invoking read.csv without specifying column names results in the first row being used as the column header. 
# b. Invoking read.csv with 'header = FALSE' specified results in the first row being automatically assigned values of "V" with increasing numbers for each column. V1, V2, V3 etc. are assigned as the names of each column.
# c. When I don't specify "na.strings = '?'", values with no data appear simply as the question mark '?' instead of 'NA'. As a result the data in those columns are treated as factors.
# d. I would imagine that the table needing to look up and assign each column a specified value rather than just with a default one (V1, V2, V3) requires more memory.
# e. Converting the data frame into a matrix basically turns all the non-numeric values into "NA". It seems like most of the numbers are still good, but all the other values become "NA" and it looks really bad.

## 4) Practice base plotting
```{r}
hist(imports_85_table$price,
     col = 'aquamarine3')

boxplot(imports_85_table$horsepower,
        horizontal = TRUE)

barplot_data <- (table(imports_85_table$body_style))
barplot(barplot_data[c(4,3,5,2,1)])

stars(imports_85_table,
      full = FALSE,
      locations = c(5,5),
      key.loc = c(5,5))
```

## 5) Summaries
```{r}
aggregate(imports_85_table[,c('price')], 
          by = list(imports_85_table$fuel_type), 
          FUN = mean, 
          na.rm = TRUE)

imports_85_table$make[(imports_85_table$num_of_cylinders == 'twelve')]

diesel_by_make <- imports_85_table$make[(imports_85_table$fuel_type == 'diesel')]
diesel_by_make_table <- table(diesel_by_make)
diesel_by_make_table[diesel_by_make_table == max(diesel_by_make_table)]

table(imports_85_table[which.max(imports_85_table$horsepower), 'price'])

quantile(imports_85_table$city_mpg, probs = 0.1)

quantile(imports_85_table$highway_mpg, probs = 0.9)

median(na.omit(imports_85_table$price[imports_85_table$city_mpg <= (quantile(imports_85_table$city_mpg, probs = 0.1))]))
```

## 6) Technical Questions about data frames
# a. When I do this I get a result that says "0 rows". The code I ran is as follows: data.frame(imports_85_table$xyz)
# b. Option 4. fails to return the vector. "mtcars[,mpg]"
# c. This command fails because when mpg is not in quotes it is looking for an object. The error message I get states that "object 'mpg' not found"
# d. YES. An R list can be used as a column because data frames are not limited to atomic vectors, they can also hold general vectors which are known as lists.
# e. Values are listed in a disorganized list that almost looks like a table but isn't really one. Each list of values has a header (such as $mpg, $disp) but the data is very hard to read as there is no uniform row indicator.
# f. I used "data.frame(abc)" and it seemed to work.

## 7) Correlations of quantitative variables
```{r}
dat <- imports_85_table[-c(1:9,15:16,18)]
qdat <- na.omit(dat)

library(corrplot)
M <- cor(qdat)
corrplot(M, method = "square")
corrplot(M, method = "square")

# I observe that there is a stong correlation between city and highway MPG. I also see that there is zero correlation between height and engine size, which is interesting. Not surprisingly, there is a strong negative correlation between both highway and city MPG and curb weight.
```

## 8.1) Principal components analysis
```{r}
prcomp_val <- prcomp(qdat,
       scale. = TRUE)
prcomp_val

eigenvalues <- prcomp_val$sdev^2
eigenvalues[1:3]
```

## 8.2) PCA plot of vehicles, and PCA plot of variables
```{r}
prcomp_val[1:2]

plot(prcomp_val)
```
